---
title: "Data analysis"
output: pdf_document
---

data input
```{r}

getwd()
enviro = read.csv("enviro.csv")
summary(enviro)
enviro <- enviro[-c(366),] #missing row at the end of data frame was removed
summary(enviro)
plot(enviro$temperature, enviro$srp)

#separate surface and hyporheic water
surf_enviro = subset.data.frame(enviro, water_type=="surface", select = -water_type)
hypo_enviro = subset.data.frame(enviro, water_type=="hyporheic", select = -water_type)

#separate by river, zone, habitat (change specs accordingly)
by_river = subset.data.frame(surf_enviro, river=="Madison", select = -river)
by_zone = subset.data.frame(surf_enviro, zone=="OC", select = -zone)
by_habitat = subset.data.frame(surf_enviro, habitat=="PN", select = -habitat)


```

tableplot package for rapid assessment
```{r}
library(tabplot)
surf_enviro2 = subset.data.frame(enviro, water_type=="surface", select = c(-water_type,-Lat,-Lon,-benthic_type))
sub<- surf_enviro2[,2:18]

tableplot(sub, fig.height=5)


```

create a dataset per river that is corrected per mean, zscores mean=0 sd=1 per floodplain.
Should modify the fancy plot to show this per river and variable
```{r}

dset = subset.data.frame(surf_enviro, river=="Swan", select = -river)
river = dset[,c(8:19)]
river2<- data.frame(matrix(0,ncol=ncol(river),nrow=nrow(river))) # or other
    for (i in 1:ncol(river)){ 
    river2[,i] <- scale(river[,i], center = TRUE, scale = TRUE)
    names(river2) <- colnames(river)
    } 
sd_SW<-river2

sd_data_temp<- rbind(sd_BT,sd_BO,sd_MA,sd_CF,sd_BH,sd_BN,sd_TN,sd_MI,sd_MK,sd_SW)
tempdata <- subset(surf_enviro, select=c(river, zone)) 
stdata<- cbind(tempdata,sd_data_temp) #contains al enviro variables standarized by floodplain
```

density plots---INTRODUCTION PLOTS? done without floodplain standarization
```{r}
library(tidyverse)
library(readr)
library(ggridges)
library(ggplot2)
library(forcats)
library(magrittr)
```
```{r}
tempdata <- subset(stdata, select=c(river, zone,nitrate)) #stdata from previous section
colnames(tempdata) <- c("river","zone", "var")
tempdata$zonemod<- ifelse(tempdata$zone=="MC", tempdata$var-2.5, tempdata$var+2.5)
tempdata$river_mod <- factor(tempdata$river, levels = c("Swan","Bitterroot","Clark fork","Boulder","Madison","Big Hole","Big Horn", "Tongue", "Missouri", "Milk"))

tempdata %>%
  ggplot(aes(y = river_mod)) +
  geom_density_ridges(aes(x = zonemod, fill = paste(river, zone)), 
           alpha = 0.8, bandwidth=0.4, 
           scale=7, color = "black", rel_min_height=0.00001,
           from = -6, to = 9) +
  labs(x = "var", y = "river", title = "plot") +
  scale_x_continuous(expand = c(0.1, 0)) +
  scale_fill_manual(values=c("#99CCFF", "#99CC66","#0066CC", "#99CC33","#99CCFF", "#99CC66","#99CCFF", "#99CC66", "#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66","#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66"))+
  theme_ridges(grid = TRUE)
```



per zone/habitat exploration

```{r}
df <- stdata # or other
 IDs<-unique(df$zone)# or other
 results<- data.frame(matrix(0,ncol=2,nrow=2)) # or other
    for (i in 1:length(IDs)){ 
    temp <- df[df$zone==IDs[i],]
    results[i,1] <- mean(temp[,3],na.rm=TRUE) # or other
    results[i,2] <- sd(temp[,3],na.rm=TRUE) # or other
    }
names<-as.data.frame(IDs)
results<-cbind(results,names)
colnames(results)<-c("zscores","sd","zone")
ggplot(results, aes(x=zone, y=zscores)) + geom_bar(stat="identity")
```

Diversity vs niche plot--need to generate a metric of volume n-dimensional. with ordination results and taking the radius as the average range of all or subset of dimensions.(selected dimensions must have similar variables with high loadings)

```{r}
library(vegan)
tempdata <- surf_enviro[,2:22]
tempdata = subset.data.frame(tempdata, select = c(-zone,-site,-habitat,-Lat,-Lon,-m,-oxygen_sat,-conductivity,-time))

df <- tempdata # or other
 IDs<-unique(df$river)# or other
 results<- data.frame(matrix(0,ncol=4,nrow=10)) # or other
    for (i in 1:length(IDs)){ 
    temp <- df[df$river==IDs[i],]
    temp<- na.omit(temp)
       temp<- as.vector(vegdist(temp[,2:12], method="mahalanobis")) #using Mahalonobis distance
       temp<-as.data.frame(temp)
         results[i,1] <- mean(temp$temp)
         results[i,2] <- sd(temp$temp)
         results[i,3] <- as.numeric(nrow(temp))
         results[i,4] <- sd(temp$temp)/sqrt(as.numeric(nrow(temp)))
    }

names<-as.data.frame(IDs)
results<-cbind(results,names)
results

```




