---
title: "Data analysis"
output: pdf_document
---

data input
```{r}

getwd()
enviro = read.csv("enviro.csv")
summary(enviro)
enviro <- enviro[-c(366),] #missing row at the end of data frame was removed
summary(enviro)
plot(enviro$temperature, enviro$srp)

#separate surface and hyporheic water
surf_enviro = subset.data.frame(enviro, water_type=="surface", select = -water_type)
hypo_enviro = subset.data.frame(enviro, water_type=="hyporheic", select = -water_type)

#separate by river, zone, habitat (change specs accordingly)
by_river = subset.data.frame(surf_enviro, river=="Madison", select = -river)
by_zone = subset.data.frame(surf_enviro, zone=="OC", select = -zone)
by_habitat = subset.data.frame(surf_enviro, habitat=="PN", select = -habitat)


```

looped groups for repetitive operations
```{r}
df <- surf_enviro # or other
 IDs<-unique(df$river)# or other
 results<- data.frame(matrix(0,ncol=1,nrow=10)) # or other
    for (i in 1:length(IDs)){ 
    temp <- df[df$river==IDs[i],]
    results[i,1] <- mean(temp[,15]) # or other
    }
names<-as.data.frame(IDs)
results<-cbind(results,names)
results

tapply(surf_enviro$doc, surf_enviro$river, mean) #if very basic operations are needed

```

create a dataset per river that is corrected per mean, zscores mean=0 sd=1 per floodplain.
Should modify the fancy plot to show this per river and variable
```{r}

dset = subset.data.frame(surf_enviro, river=="Clark fork", select = -river)
river = dset[,c(8:19)]
river2<- data.frame(matrix(0,ncol=ncol(river),nrow=nrow(river))) # or other
    for (i in 1:ncol(river)){ 
    river2[,i] <- scale(river[,i], center = TRUE, scale = TRUE)
    names(river2) <- colnames(river)
    } 
sd_CF<-river2

sd_data_temp<- rbind(sd_BT,sd_BO,sd_MA,sd_CF,sd_BH,sd_BN,sd_TN,sd_MI,sd_MK,sd_SW)
tempdata <- subset(surf_enviro, select=c(river, zone)) 
stdata<- cbind(tempdata,sd_data_temp) #contains al enviro variables standarized by floodplain
```

density plots---INTRODUCTION PLOTS? done without floodplain standarization
```{r}
library(tidyverse)
library(readr)
library(ggjoy)
library(ggplot2)
```
```{r}
stdata<-as.data.frame(stdata)
tempdata <- subset(stdata, select=c(river, zone,nitrate)) 
colnames(tempdata) <- c("river","zone", "var")
tempdata$zonemod<- ifelse(tempdata$zone=="MC", (tempdata$var*-1)-10, (tempdata$var*1)+10)

 sapply(tempdata, class)



ggplot(tempdata, aes(x = zonemod, y = river, group=river)) +
        geom_joy_gradient(data=filter(tempdata, zone=="MC"), scale = 8, size = 0.15, rel_min_height = 0.01, fill="#0066CC", alpha=0.9,bandwidth=4) +
        geom_joy_gradient(data=filter(tempdata, zone=="OC"), scale = 8, size = 0.15, rel_min_height = 0.01, fill="#99CC33", alpha=0.9,bandwidth=4) +
        theme_joy() +
        scale_x_continuous(limits=c(-50, 50),expand = c(0.01, 0)) 
       
```












Diversity vs niche plot--need to generate a metric of volume n-dimensional. with ordination results and taking the radius as the average range of all or subset of dimensions.(selected dimensions must have similar variables with high loadings)

```{r}
library(vegan)
tempdata <- surf_enviro[,2:22]
tempdata = subset.data.frame(tempdata, select = c(-zone,-site,-habitat,-Lat,-Lon,-m,-oxygen_sat,-conductivity,-time))

df <- tempdata # or other
 IDs<-unique(df$river)# or other
 results<- data.frame(matrix(0,ncol=4,nrow=10)) # or other
    for (i in 1:length(IDs)){ 
    temp <- df[df$river==IDs[i],]
    temp<- na.omit(temp)
       temp<- as.vector(vegdist(temp[,2:12], method="mahalanobis")) #using Mahalonobis distance
       temp<-as.data.frame(temp)
         results[i,1] <- mean(temp$temp)
         results[i,2] <- sd(temp$temp)
         results[i,3] <- as.numeric(nrow(temp))
         results[i,4] <- sd(temp$temp)/sqrt(as.numeric(nrow(temp)))
    }

names<-as.data.frame(IDs)
results<-cbind(results,names)
results

```




