---
title: "Data analysis"
output: pdf_document
---

data input
```{r}

getwd()
enviro = read.csv("enviro.csv")
summary(enviro)
enviro <- enviro[-c(366),] #missing row at the end of data frame was removed
summary(enviro)
plot(enviro$temperature, enviro$srp)

#separate surface and hyporheic water
surf_enviro = subset.data.frame(enviro, water_type=="surface", select = -water_type)
hypo_enviro = subset.data.frame(enviro, water_type=="hyporheic", select = -water_type)

```

create a dataset per river that is corrected per mean, zscores mean=0 sd=1 per floodplain.
```{r}
rivers <- c("Swan","Bitterroot","Clark fork","Boulder","Madison","Big Hole","Big Horn", "Tongue", "Missouri", "Milk")
    
df_list <- split(surf_enviro[,9:20], as.factor(surf_enviro$river))
df_list2 <- matrix(0,ncol=12,nrow=0)
    for (i in 1:length(df_list)){ 
    temp <- scale(df_list[[i]], center = TRUE, scale = TRUE)
    df_list2<-rbind(df_list2,temp)
    } 
df_list2<-as.data.frame(df_list2) 
st_data<-cbind(surf_enviro[,1:5],df_list2)
#contains all rivers and  enviro variables standarized 


```

Diversity vs niche plot--need to generate a metric of volume n-dimensional. with ordination results and taking the radius as the average range of all or subset of dimensions.(selected dimensions must have similar variables with high loadings)

```{r}
library(vegan)
pos <- c(5,15,20,10,25,30,35,40,45,0)
tempdata = subset.data.frame(st_data, select = c(river,zone,oxygen_sat,temperature,doc,ammonium,nitrate,phosphate,tn,tp))
tempdata = tempdata[complete.cases(tempdata), ]
df <- tempdata # or other
 IDs<-unique(df$river)# or other
 results<- data.frame()
    for (i in 1:length(IDs)){ 
    temp <- df[df$river==IDs[i],]
    temp_MC<- subset.data.frame(temp,zone=="MC",select = c(-river,-zone))
    temp_OC<- subset.data.frame(temp,zone=="OC",select = c(-river,-zone))
      temp_MC<- as.vector(vegdist(temp_MC, method="euclidean",na.rm=TRUE))
      temp_OC<- as.vector(vegdist(temp_OC, method="euclidean",na.rm=TRUE))
        temp_MC<-as.data.frame(temp_MC)
        riv<-as.data.frame(rep(IDs[i], length(temp_MC$temp_MC)))
        zon<-as.data.frame(rep("MC", length(temp_MC$temp_MC)))
        temp_MC<-cbind(riv,zon,temp_MC)
              colnames(temp_MC)<- c("river","zone","var")
              temp_MC$var<-temp_MC$var+pos[i]
        temp_OC<-as.data.frame(temp_OC)
        riv<-as.data.frame(rep(IDs[i], length(temp_OC$temp_OC)))
        zon<-as.data.frame(rep("OC", length(temp_OC$temp_OC)))
        temp_OC<-cbind(riv,zon,temp_OC)
              colnames(temp_OC)<- c("river","zone","var")
              temp_OC$var<-temp_OC$var+pos[i]
         results<-rbind(results,temp_MC)
         results<-rbind(results,temp_OC)
            temp_MC <-NULL
            temp_OC <-NULL
    }

```


density plots--- done after floodplain standarization and niche aproximation
```{r}
library(tidyverse)
library(readr)
library(ggridges)
library(ggplot2)
library(forcats)
library(magrittr)
```
```{r}
tempdata <- results
tempdata$zonemod<- ifelse(tempdata$zone=="MC", tempdata$var-0, tempdata$var+0)
tempdata$river_mod <- factor(tempdata$river, levels = c("Swan","Bitterroot","Clark fork","Boulder","Madison","Big Hole","Big Horn", "Tongue", "Missouri", "Milk"))

tempdata %>%
  ggplot(aes(y = river_mod)) +
  geom_density_ridges(aes(x = zonemod, fill = paste(river, zone)), 
           alpha = 0.8, bandwidth=1, 
           scale=3, color = "black", rel_min_height=0.00001,
           from = -5, to = 55) +
  labs(x = "var", y = "river", title = "Euclidean distances of scaled environmental variables") +
  scale_x_continuous(expand = c(0.1, 0)) +
  scale_fill_manual(values=c("#99CCFF", "#99CC66","#0066CC", "#99CC33","#99CCFF", "#99CC66","#99CCFF", "#99CC66", "#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66","#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66"))+
  theme_ridges(grid = TRUE)+
  theme(legend.position="none")
```


