---
title: "Data analysis"
output: pdf_document
---
using package 'hydrostats' te get baseflow and colwells statistics for the 1970-2010 period for each river system---used in Figure 1
```{r}
library(hydrostats)
```
```{r}
data <-read.table("clipboard", header = T, sep = "\t")
colnames(data)<-c("Date","Q")
data<-ts.format(data)
Colwells(data)
out<- baseflows(data, ts = "annual")
mean(0.0283168*out$bf)
sd(0.0283168*out$bf)
```
```{r}
data <-read.table("clipboard", header = T, sep = "\t")
colnames(data)<-c("Date","Q")
data<-ts.format(data)
out<- Colwells(data)
out

data(Acheron)
Acheron<-ts.format(Acheron)
                   
```

data input
```{r}
getwd()
enviro = read.csv("enviro.csv")
summary(enviro)
plot(enviro$temperature, enviro$srp)
surf_enviro=enviro
```

create a dataset per river that is corrected per mean, zscores mean=0 sd=1 per floodplain.
```{r}
df_list <- split(surf_enviro[,7:18], as.factor(surf_enviro$river))
df_list2 <- matrix(0,ncol=12,nrow=0)
    for (i in 1:length(df_list)){ 
    temp <- scale(df_list[[i]], center = TRUE, scale = TRUE)
    df_list2<-rbind(df_list2,temp)
    } 
df_list2<-as.data.frame(df_list2) 
st_data<-cbind(surf_enviro[,2],df_list2)
names(st_data)[1]<-paste("zone")
#contains all rivers and  enviro variables standarized 
```
```{r}
#averaging by habitat--
library(doBy)
habitat_data_avg<-summaryBy(doc+ammonium+nitrate+phosphate+tn+tp ~ habitat, data = st_data,
  FUN = function(x) { c(m = mean(x,na.rm = TRUE)) } )
```
```{r}
#averaging by zone--
library(doBy)
zone_data_avg<-summaryBy(doc+ammonium+nitrate+phosphate+tn+tp ~ zone, data = st_data,
  FUN = function(x) { c(m = mean(x,na.rm = TRUE)) } )
```
```{r}
#averaging by river--
library(doBy)
enviro_river<-surf_enviro[,7:18]
enviro_river<-as.data.frame(scale(enviro_river, center = TRUE, scale = TRUE))
enviro_river<-cbind(surf_enviro[,1],enviro_river)
names(enviro_river)[1]<-paste("river")
river_data_avg<-summaryBy(doc+ammonium+nitrate+phosphate+tn+tp ~ river, data = enviro_river,
  FUN = function(x) { c(m = mean(x,na.rm = TRUE)) } )
```

Radius plots or habitat and river enviro variation
```{r}
library(fmsb)
row.names(river_data_avg)<-river_data_avg[,1]
river_data_avg[,1] <-NULL
river_data_avg=rbind(rep(3,7) , rep(-1,7) , river_data_avg)
#plot
colors_border=c("#FF3300","#CCFF33","#CCCC00","#FF6633",
                "#3399FF","#000033","#FF9933","#CC66CC",
                "#00CCFF","#999999")
radarchart( river_data_avg  , axistype=1 ,
    #custom polygon
    pcol=colors_border , pfcol=NULL , plwd=4 , plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(-1,3,1), cglwd=0.8,
    #custom labels
    vlcex=0.8
    )
legend(x=1.2, y=1, legend = rownames(river_data_avg[-c(1,2),]), bty = "n", pch=20 , col=colors_border , text.col = "grey", cex=1.2, pt.cex=3)
```





Diversity vs niche plot--need to generate a metric of volume n-dimensional. 

```{r}
library(vegan)
pos <- c(5,15,20,10,25,30,35,40,45,0)
tempdata = subset.data.frame(st_data, select = c(river,zone,oxygen_sat,temperature,doc,ammonium,nitrate,phosphate,tn,tp))
tempdata = tempdata[complete.cases(tempdata), ]
df <- tempdata # or other
 IDs<-unique(df$river)# or other
 results<- data.frame()
    for (i in 1:length(IDs)){ 
    temp <- df[df$river==IDs[i],]
    temp_MC<- subset.data.frame(temp,zone=="MC",select = c(-river,-zone))
    temp_OC<- subset.data.frame(temp,zone=="OC",select = c(-river,-zone))
      temp_MC<- as.vector(vegdist(temp_MC, method="euclidean",na.rm=TRUE))
      temp_OC<- as.vector(vegdist(temp_OC, method="euclidean",na.rm=TRUE))
        temp_MC<-as.data.frame(temp_MC)
        riv<-as.data.frame(rep(IDs[i], length(temp_MC$temp_MC)))
        zon<-as.data.frame(rep("MC", length(temp_MC$temp_MC)))
        temp_MC<-cbind(riv,zon,temp_MC)
              colnames(temp_MC)<- c("river","zone","var")
              temp_MC$var<-temp_MC$var+pos[i]
        temp_OC<-as.data.frame(temp_OC)
        riv<-as.data.frame(rep(IDs[i], length(temp_OC$temp_OC)))
        zon<-as.data.frame(rep("OC", length(temp_OC$temp_OC)))
        temp_OC<-cbind(riv,zon,temp_OC)
              colnames(temp_OC)<- c("river","zone","var")
              temp_OC$var<-temp_OC$var+pos[i]
         results<-rbind(results,temp_MC)
         results<-rbind(results,temp_OC)
            temp_MC <-NULL
            temp_OC <-NULL
    }

```


density plots---niche dimension representation per floodplain with zone diferentiation
```{r}
library(tidyverse)
library(readr)
library(ggridges)
library(ggplot2)
library(forcats)
library(magrittr)
```
```{r}
tempdata <- results
tempdata$zonemod<- ifelse(tempdata$zone=="MC", tempdata$var-0, tempdata$var+0)
tempdata$river_mod <- factor(tempdata$river, levels = c("Swan","Bitterroot","Clark fork","Boulder","Madison","Big Hole","Big Horn", "Tongue", "Missouri", "Milk"))

tempdata %>%
  ggplot(aes(y = river_mod)) +
  geom_density_ridges(aes(x = zonemod, fill = paste(river, zone)), 
           alpha = 0.8, bandwidth=1, 
           scale=3, color = "black", rel_min_height=0.00001,
           from = -5, to = 55) +
  labs(x = "var", y = "river", title = "Euclidean distances of scaled environmental variables") +
  scale_x_continuous(expand = c(0.1, 0)) +
  scale_fill_manual(values=c("#99CCFF", "#99CC66","#0066CC", "#99CC33","#99CCFF", "#99CC66","#99CCFF", "#99CC66", "#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66","#0066CC", "#99CC33","#0066CC", "#99CC33","#99CCFF", "#99CC66"))+
  theme_ridges(grid = TRUE)+
  theme(legend.position="none")
```


